Ladies' figure skating in the 2002 Winter Olympics

### 2022-10-18 inst/skating.Rmd
loads of fixes, accounts for name ordering this time

```{r,maxshow, cache = TRUE}
skating <- ordertable2supp(skating_table)
skating_maxp <- maxp(skating)
m <- skating_maxp # for ease of  typing
```

@@

```{r,longthing,cache=TRUE}
jj <-
    matrix(c(
        -1, 0, 0,   1,0,  # Hughes   < Cohen
         0,-1, 0,   1,0,  # Slutskya < Cohen
         0, 0,-1,   1,0,  # Kwan     < Cohen

        -1, 0, 0,   0,1,  # Hughes   < Suguri
         0,-1, 0,   0,1,  # Slutskya < Suguri
         0, 0,-1,   0,1), # Kwan     < Suguri
        byrow=TRUE,ncol=5)

problem_constraints <-  # fill with zeros for other competitors
    cbind(jj,matrix(0,nrow(jj),size(skating)-ncol(jj)-1))

small <- 1e-4  # need a sensible start value satisfying the constraints
start <- c(rep(2*small,3),rep(3*small,2),rep(small,17))

@@

for(i in seq_len(nrow(problem_constraints))){
  jj <- maxp(skating, startp=start, give=TRUE,fcm=problem_constraints[i,], fcv=0,n=1,SMALL=1e-5)
   fullout[[i]] <- jj
   out[i] <- jj$value
}
out
```

@@

```{r,unconstrainedlikes, cache=TRUE}
mgv <- maxp(skating, give=TRUE, n=10)$value
mgv  - out
mgv  - max(out)
```

that is, a little over 1 unit of support, falling short of the two
units suggested by Edwards.  Observe that the maximum likelihood among
the six alternative hypotheses is that of number 3, in which the
maximization was constrained to obey `Kwan < Cohen`.

If, instead, we ask whether there is evidence that Suguri should not
have been a medallist, we find

@@

We can be more specific with the likelihoods:

>>>>

```{r,maxshow, cache = TRUE}
skating <- ordertable2supp(skating_table)
skating_maxp <- maxp(skating)
m <- skating_maxp # for ease of  typing
m
```

```{r equalskatingptest}
equalp.test(skating)
```

@@

```{r}
f <- function(smaller,bigger){
  jj <- rep(0,22)
  names(jj) <- pnames(skating)[-23]  #  no fillup
  jj[smaller] <- -1
  jj[bigger ] <- +1
  return(jj)
}

problem_constraints <-  NULL
problem_constraints %<>% rbind(f("hughes"   ,"cohen" ))
problem_constraints %<>% rbind(f("slutskaya","cohen" ))
problem_constraints %<>% rbind(f("kwan"     ,"cohen" ))
problem_constraints %<>% rbind(f("hughes"   ,"suguri"))
problem_constraints %<>% rbind(f("slutskaya","suguri"))
problem_constraints %<>% rbind(f("kwan"     ,"suguri"))

rownames(problem_constraints) <-
c("hug < coh", "slu < coh", "kwa < coh", "hug < sug", "slu < sug", "kwa < sug" )
colnames(problem_constraints) %<>% substr(1,3)
problem_constraints

small <- 0.01
big <- 0.1
start_vector <- rep(small,22)
names(start_vector) <- pnames(skating)[-23]  #  no fillup
start_vector["cohen"] <- big
start_vector["suguri"] <- big
```

```{r,longthing,cache=TRUE}

@@

mle_skaters <- NULL

for(i in seq_len(nrow(problem_constraints))){
  jj <- maxp(skating, startp=start_vector, give=TRUE,fcm=problem_constraints[i,], fcv=0,n=10,SMALL=1e-5)
   fullout[[i]] <- jj
   out[i] <- jj$value
   mle_skaters <- cbind(mle_skaters,jj$par)

}
out
out-max(out)
colnames(mle_skaters) <- rownames(problem_constraints)
mle_skaters
```

@@

```{r,unconstrainedlikes, cache=TRUE}
mgv <- maxp(skating, give=TRUE, n=10)$value
mgv
mgv  - out
mgv  - max(out)
```

that is, a little over 7 units of support, clearly exceeding the two
units suggested by Edwards.  We have strong evidence to suggest that
the medallists were indeed the strongest three competitors.  Observe
that the maximum likelihood among the six alternative hypotheses is
that of number 3, in which the maximization was constrained to obey
`Kwan < Cohen`.  If we _have_ to change the medallists, then
exchanging the order of Kwan and Cohen incurs the lowest likelihood
penalty.  If, instead, we ask whether there is evidence that Suguri
[who was much weaker than Cohen] should not have been a medallist, we
find

@@

Much stronger.
